\subsubsection{Data Contamination}

We observed significant performance gains in \ac{SoTA} \acp{LLM} across the
board, which might point to potential data contamination, i.e., the models have
seen some of the clue-answer pairs during training. To ensure that our observed 
performance enhancements were not merely a result of data contamination, we 
assembled additional datasets consisting of cryptic crossword clues, all sourced 
from puzzles released after \emph{May 20, 2024, which is after the knowledge 
cutoff for all \acp{LLM} evaluated} (see Table \ref{tab:dataset_details}). To 
guard against any inadvertent duplications, we checked the answers in these 
\emph{post-cutoff} datasets against the entire pool of cryptic crossword 
datasets utilized within our study, totaling 665,497 answers. We found no 
duplicates in the \emph{post-cutoff} Guardian set and only two in the 
\emph{post-cutoff} Lovatts set, which were removed. The \emph{Init} dataset is 
also derived from \emph{The Guardian}, ensuring that the results displayed in
Table \ref{tab:post_cutoff} maintain consistency.

\begin{table}[!ht]
    \caption{
        \textbf{Performance of LLMs on \emph{post-cutoff} datasets.}\\
        Note, \emph{Init} by \citet{curriculum}, also sourced their data from 
        \emph{The Guardian}, thus these results provide a fair head-to-head 
        comparison of performance. We report exact match $(\%)$.
    }
    \label{tab:post_cutoff}
    \begin{center}
        \begin{tabular}[c]{ l r r r }
            \toprule
            Model           & \textbf{Lovatts}    & \textbf{Guardian} & \textbf{Init} \\
            \midrule
            Llama 3 70B     $\qquad$&    26.03\%     & 5.5 \% & 6.4 \%  \\
            Claude 3 Sonnet $\qquad$&    46.28\%     & 12.5\% & 10.8\%  \\
            GPT 4 Turbo     $\qquad$&    61.57\%     & 18.5\% & 18.7\%  \\
            \bottomrule
        \end{tabular}
    \end{center}
\end{table}

Since \emph{no appreciable difference in performance on the post-cutoff dataset}
(see Table \ref{tab:post_cutoff}) is observed, we conclude that these \acp{LLM}
can generalize beyond potential contamination in their training set.

\subsubsection{Human Evaluation}

To determine the capacity of models to reason about cryptic crossword clues, we
perform \textbf{human evaluation}. We used a 3-shot \emph{Chain-of-thought} 
prompt to elicit a reasoned response to crossword clues from GPT-4-Turbo, which 
are then analyzed for soundness vis-\`a-vis factual and logical errors. In 
checking whether a justification for an answer given by the \ac{LLM} is 
logically and factually sound, we assess grammatical soundness and phraseological 
meaningfulness of the sentences in the answer, the existence of counterfactual 
statements (e.g., ``BULKY has 4 characters'', ``the initial letters of ARE RATS 
TIRED NOW are ARTS'', etc.), and whether it presents a statement as an inference 
from previous statements when it does not follow from those, etc. If an answer 
by the \ac{LLM} is found unsatisfactory in any of these aforementioned areas, it 
is labeled \textbf{unsound}. All responses were evaluated by 3 annotators\footnote{
authors of \citet{myCrossword}.} and in case of conflicting answers (4 out of 
100, Fleiss' $\kappa$ = 0.94), discussions were held to reach a consensus. We 
chose the 100 samples from the post-cutoff \emph{Lovatts} set for this to allay 
concerns of contamination. 

\begin{table}[!ht]
    \caption{
        \textbf{Results from human evaluation.}\\
        An answer is called \textbf{correct} if the model prediction exactly 
        matches the ground truth. The answer is called sound if it contains no 
        logical or factual errors. Results are with GPT-4-Turbo on the 
        \emph{post-cutoff} Lovatts set.
    }
    \label{tab:human-eval}
    \begin{center}
        \begin{tabular}[c]{ l|l|r }
            \toprule
            \multicolumn{1}{c}{}  & \multicolumn{1}{c}{\textbf{Sound}}   & \multicolumn{1}{c}{$\neg$\textbf{Sound}} \\[4pt]
            \multicolumn{1}{c}{}  & \multicolumn{1}{c}{(61)}             & \multicolumn{1}{c}{(39)}                  \\
            \midrule
            Correct (65) $\qquad$ &  48\%    & 17\% \\
            \cline{2-3}
            Wrong (35)   $\qquad$ &  13\%    & 22\% \\
            \bottomrule
        \end{tabular}
    \end{center}
\end{table}

The results (see Table \ref{tab:human-eval}) show that \textbf{74\%} of the time 
GPT-4-Turbo provided a correct answer, it also gave \emph{sound reasoning in 
support of the answer}. This leads us to conjecture that they possibly have a
significant ability to reason and generalize.

\subsubsection{Potential Pitfalls}

To further analyze if \acp{LLM} demonstrate any common failure modes, we 
manually tagged the human evaluation dataset based on the principal skill 
required to solve a particular puzzle clue. The skill-based categories are:
\begin{itemize}
    \item \textbf{ANG}--the answer is an \emph{anagram} of some words of the
        clue (e.g., \texttt{\colorbox{shadecol}{Cubit is} mixed up cookie } $\rightarrow$ \texttt{BISCUIT}).
    \item \textbf{HOM}--the answer is a \emph{homophone} of some words of the 
        clue (e.g., \texttt{Heard \colorbox{shadecol}{prints} are for royalty} $\rightarrow$ \texttt{PRINCE}).
    \item \textbf{CNT}--the answer is disguised in a contiguous section of the 
        clue (e.g., \texttt{The Press leaves pres\colorbox{shadecol}{enter}s to go in} $\rightarrow$ \texttt{ENTER}).
    \item \textbf{SCJ}--the answer is found by combining several words that are
        synonyms of various parts of the clue (e.g., \texttt{Reasonable \colorbox{boxcol}{food allowance} for \colorbox{shadecol}{Capone}} $\rightarrow$ \texttt{\colorbox{boxcol}{RATION}\colorbox{shadecol}{AL}}).
    \item \textbf{OTH}--this class lumps together a variety of other skills like
        Spoonerisms, acronyms, world knowledge, and various other kinds of
        character manipulations 
        (e.g., \texttt{\colorbox{shadecol}{Pi}p\colorbox{shadecol}{er} loses
        heart on jetty} $\rightarrow$ \texttt{PIER}).
\end{itemize}

\begin{table}[!ht]
    \caption{
        \textbf{Are there common failure modes for \acp{LLM}?}\\
        \textbf{ANG} refers to \emph{anagrams}, \textbf{SCJ} refers to 
        \emph{synonym conjugation}, \textbf{CNT} refers to \emph{containment}, 
        \textbf{HOM} refers to \emph{homophones}, and \textbf{OTH} refers to 
        \emph{others}. The percentages in parentheses refer to the prevalence
        of a particular type of clue in the database.
    }
    \label{tab:failure-mode}
    \begin{center}
        \begin{subtable}{\textwidth}
            \centering
            \begin{tabular}[c]{lrrrrr}
                \toprule
                \multirow{2}{*}{\emph{Lovatts}} & \textbf{ANG}   & \textbf{CNT}      & \textbf{SCJ}  & \textbf{HOM}  & \textbf{OTH}  \\
                                                & (27\%)         & (26\%)            & (18\%)        & (8\%)         & (21\%)        \\
                \midrule
                GPT 4 Turbo     & 74\%           & 56\%              & 50\%          & 100\%         & 67\% \\
                \bottomrule
            \end{tabular}
            \caption{Results for \emph{Lovatts} dataset (human-annotated).}
            \label{tab:failure-mode1}
        \end{subtable}

        \begin{subtable}{.45\textwidth}
            \centering
            \begin{tabular}[c]{lrr}
                \toprule
                \multirow{2}{*}{\emph{Init}+\emph{Cryptonite}} & \textbf{ANG} & \textbf{CNT} \\
                                                    &   (8.5\%)      & (2.5\%)               \\
                \midrule
                Llama 3 70B     & 2.6\%        & 12.2\%                                          \\
                Claude 3 Sonnet & 7.9\%        & 15.3\%                                          \\
                GPT 4 Turbo     & 25.0\%       & 33.7\%                                          \\
                \bottomrule
            \end{tabular}
            \caption{Results for \emph{Init}+\emph{Cryptonite} dataset.}
            \label{tab:failure-mode2}
        \end{subtable}
    \end{center}
\end{table}

Due to the limited number of human annotations, our results (Table 
\ref{tab:failure-mode1}) are not statistically significant; however, these 
results hint at the fact that GPT-4-Turbo demonstrates strong performance in 
\emph{anagrams} and \emph{homophone}-based clues, boasting 74\%  and 100\% 
accuracy, respectively (baseline\footnote{Baseline refers to mean accuracy across 
all kinds of clues.} - 65\%).

We extended this analysis to the \emph{Cryptonite} and \emph{Init} datasets (Table \ref{tab:failure-mode2}),
which together include 4000 clue-answer pairs, finding similar trends:
GPT-4-Turbo maintained respectable anagram performance with 25\% accuracy over
a 21.1\% baseline. Conversely, Llama 3 70B recorded a meager 2.6\% accuracy,
lagging behind a 7.16\% baseline, and Claude 3 Sonnet mirrored this trend with
7.9\% accuracy compared to an 11.85\% baseline. These findings suggest that
Llama 3 70B and Claude 3 Sonnet may struggle more with anagram-based clues. It
should be noted that unlike \textbf{(ANG, CNT)}, categories like \textbf{(HOM,
SCJ, OTH)} cannot be reliably identified automatically, presenting a limitation
in broader analysis for these types of clues.

To better understand the influence of \emph{sub-token counting performance} on 
clue-solving capabilities, we crafted an additional experiment. We consider all
such clues where the model successfully interpreted the clue’s semantics but 
failed to comply with the specified length constraints (e.g., \texttt{LECTURER} 
instead of \texttt{PROFESSOR} or \texttt{NANNA} instead of \texttt{GRANNY}). We
counted the number of \emph{wrong \ac{LLM} predictions with high semantic 
similarity to ground truth answers}.\footnote{Similarity score of 0.5 or higher 
as measured by OpenAI \texttt{text-embedding-3-large}} We found that 
GPT-4-Turbo and Llama 3 70B, respectively, produced length error predictions 
\textbf{46.4\%} and \textbf{59.9\%} of the time. These findings underscore how 
weak length constraint adherence ability severely hampers the clue-solving 
potential of \acp{LLM}, highlighting a key area for improvement.
