\Acf{DL} refers to a powerful and diverse family of techniques that leverage 
data to model complex real world phenomena. What sets apart \ac{DL} from the
majority of other techniques in this regime is its unparalleled versatility.
\Ac{DL} has been successfully applied to, and attained \ac{SoTA} results 
in, domains like \ac{CV}, \ac{NLP}, reinforcement learning, biotechnology 
\cite{alphafold}, medicine \cite{ng}, and hundreds more.

Although \ac{DL} features multifarious techniques, architectures, training 
paradigms, and methodologies, there are a few unifying factors. First and most
important of which is \emph{data}, and lots of it. This data can be un-annotated
\emph{(unsupervised)} or annotated by crowdsourcing or appointing experts 
\emph{(supervised)}. Also, available data is often cleverly repurposed by making 
assumptions or employing additional techniques \emph{(weakly supervised)}. 
Secondly, \ac{DL}-based techniques feature at least one parametrized non-linear 
almost-everywhere differentiable function (\emph{model}), which is meant to
imitate the underlying relationship we wish to explore. Finally, they feature an
objective or \emph{loss} function, quantifying desired outcomes for the model
when applied to the dataset in question. With just these three ingredients, a
large array of problems in varied areas can be tackled.

In this section, we go over some of the background necessary for the rest of the 
presentation. We do not provide a thorough overview of \ac{DL}, and we recommend
the work by \citet{goodfellowDL} for an expansive treatment.
