\thispagestyle{plain}
\newgeometry{left=30mm,right=30mm,top=0mm,bottom=30mm,heightrounded=true}
\onehalfspacing
\chapter*{Acknowledgements} % (fold)
\addcontentsline{toc}{chapter}{Acknowledgements}
\label{chap:Acknowledgements}


Embarking on this doctoral journey, I truly felt like a fish out of water, but
over time I discovered a fascinating world full of remarkable possibilities that
captured my imagination and fueled my passion for inquiry. Completing this 
journey has been profoundly rewarding, a milestone made possible by the 
unwavering support and encouragement of many remarkable individuals. 

First and foremost, I would like to extend my heartfelt gratitude to 
my advisor, Prof. Utpal Garain. Your invaluable guidance, expertise, and 
patience have been instrumental in every facet of my research. Your belief in 
my capabilities has been a profound source of motivation. I am incredibly 
fortunate to have had a mentor of your caliber.

I am forever indebted to my parents, whose unconditional love and steadfast
encouragement have been a bedrock of support throughout my academic endeavors. 
Your sacrifices and faith in my potential have been the driving force behind my 
pursuit of this academic milestone. Thank you for always believing in me.

I extend a special thanks to Akshay Chaturvedi for showing me the 
ropes. I am also grateful to my colleagues and friends - Joy Mahapatra,
Saptarshi Saha, and Aditya Shankar Pal who have offered their support and 
encouragement throughout this process. I have been immensely blessed with the 
opportunity to collaborate with exceptional individuals, and I extend my sincere
thanks to Nicholas Asher, Supratik Pal, Rahul Shah, Purba Mukherjee, Arijit 
Ukil, Arpan Pal, Claudia Stern, Scott Ritter, and Aude Benk-Fortin for being a 
part of this journey and for their valuable insights.

I am deeply grateful to my institute for providing this opportunity to pursue my
doctoral studies, and for fostering a community of nurturing individuals who
have played a crucial and supportive role in my academic journey. I cherish the
memories and experiences gained within this esteemed institution.

Lastly to my partner, Sutanoya, your understanding, affection, and unwavering 
support have been my anchor during the highs and lows of this journey. Having 
an in-house world-class mathematician and newspaper-worthy copy editor is an 
incredibly valuable resource, especially when it only comes at the cost of an 
equal share of my caloric intake. Jokes aside, your presence has been a source 
of strength and comfort, making this journey all the more meaningful.

\hfill\LARGE{$\blacksquare$}\normalsize

\newpage
\null\thispagestyle{empty}
\newpage
% chapter Acknowledgements (end)
