\makeatletter\@openrightfalse
\cleardoublepage
\chapter*{Publications}
\addcontentsline{toc}{chapter}{Publications}

Following is a list of publications pertinent to this dissertation.
\begin{spacing}{0.9}
\small{
\begin{itemize}
    \item \emph{MedTric : A clinically applicable metric for evaluation of multi-label computational diagnostic systems}.
        \hfill\colorbox{shadecol}{\emph{Journal}}\\
        \textbf{Soumadeep Saha}, Utpal Garain, Arijit Ukil, Arpan Pal, Sundeep Khandelwal.\\
        PLOS ONE, 18(8), 1--19; 2023. [\texttt{10.1371/journal.pone.0283895}] \cite{myMedtric}.
    \item \emph{Analyzing Semantic Faithfulness of Language Models via Input Intervention on Question Answering}.
        \hfill\colorbox{shadecol}{\emph{Journal}}\\
        Akshay Chaturvedi, Swarnadeep Bhar, \textbf{Soumadeep Saha}, Utpal Garain, Nicholas Asher.\\
        Computational Linguistics, 50(1), 119-155; 2024. [\texttt{10.1162/coli\_a\_00493}] \cite{myCL}.
    \item \emph{{VALUED} - Vision and Logical Understanding Evaluation Dataset}.
        \hfill\colorbox{shadecol}{\emph{Journal}}\\
        \textbf{Soumadeep Saha}, Saptarshi Saha, Utpal Garain.\\
        Journal of Data-centric Machine Learning Research (DMLR), (13):1--18; 2024. MLR press \cite{myValued}.
    \item \emph{LADDER: Revisiting the Cosmic Distance Ladder with Deep Learning Approaches and Exploring Its Applications}.
        \hfill\colorbox{shadecol}{\emph{Journal}}\\
        Rahul Shah, \textbf{Soumadeep Saha}, Purba Mukherjee, Utpal Garain, Supratik Pal.\\
        The Astrophysical Journal Supplement Series (ApJS), 273(2), 1--27; 2024. The American Astronomical Society. [\texttt{10.3847/1538-4365/ad5558}] \cite{myLadder}.
    \item \emph{Language Models are Crossword Solvers}.
        \hfill\colorbox{shadecol}{\emph{Conference}}\\
        \textbf{Soumadeep Saha}, Sutanoya Chakraborty, Saptarshi Saha, Utpal Garain.\\
        2025 Annual Conference of the Nations of the Americas Chapter of the Association for Computational Linguistics [NAACL 2025 (main)], 2074--2090; 2025. Association for Computational Linguistics. [\texttt{10.18653/v1/2025.naacl-long.104}] \cite{myCrossword}.
    \item \emph{DOST--Domain Obedient Self-supervision for Trustworthy Multi Label Classification with Noisy Labels}.
        \hfill\colorbox{shadecol}{\emph{Workshop}}\\
        \textbf{Soumadeep Saha}, Utpal Garain, Arijit Ukil, Arpan Pal, Sundeep Khandelwal.\\
        Proceedings of the 8th International Workshop in Health Intelligence, The Association for the Advancement of Artificial Intelligence (AAAI), 2024.\\
        AI for Health Equity and Fairness: Leveraging AI to Address Social Determinants of Health, 117--127; 2024. Springer Nature Switzerland. [\texttt{10.1007/978-3-031-63592-2\_10}] \cite{myDost}.
\end{itemize}}
\end{spacing}

\newpage\null\thispagestyle{empty}\newpage
\@openrighttrue\makeatother 
