The dissertation starts (Chapter 1) with a summary of requisite background
pertaining to \ac{DL} techniques, architectures, language modeling, and related
prior work in this area. This is followed by \emph{four} contributory chapters.

\textbf{Chapter 2} \emph{(Do vision systems learn rules?)} focuses
on \ac{DL} systems' ability to adhere to domain rules for vision tasks. It
introduces a new dataset to combat the lack of large-scale rule-annotated
datasets and, with its assistance, demonstrates the constraint adherence deficit
displayed by \ac{SoTA} \ac{DL}-based \ac{CV} techniques. Further analysis
demonstrates key limitations of vision models in performing crucial tasks like
counting and localization.

\textbf{Chapter 3} \emph{(Faithful Language Modeling)} analyzes \acp{LM} and
points out critical deficiencies they exhibit in adhering to domain expectations
with the aid of interventions during training and inference. A notion of
semantic faithfulness is introduced, which demands that \acp{LM}' answers should
change in response to withholding of relevant context, and it is empirically
demonstrated that \acp{LM} do not abide by this requirement. Further, an
intervention-based training strategy is proposed that alleviates this effect.

\textbf{Chapter 4} \emph{(Domain-aware Learning and Evaluation)} puts forth a
technique called Domain Obedient Self-supervised Training (DOST) to incorporate
logical constraint information into \ac{DL} systems. DOST leverages domain rules
alongside data to disincentivize incoherent predictions and improve predictive
performance. This algorithm is more data-efficient and results in models that
show enhanced \emph{trustworthiness}. Additionally, chapter 4 tackles model
evaluation and points out issues with a domain-blind approach to evaluation. A
framework for constructing a metric that takes domain knowledge into account is
proposed and exemplified with a real-life medical use case.

\textbf{Chapter 5} \emph{(Constrained Inference)} explores inference with
\acp{LLM} in a constrained setting. In particular, crosswords--a word puzzle
featuring rich linguistic requirements alongside constraint satisfaction--are
analyzed. Various subtasks of this problem are examined with an emphasis on
isolating key areas of weaknesses for \acp{LLM}. Further, a new algorithm is
presented, which uses in-context learning paired with a search strategy to
successfully solve crosswords. This chapter presents several results that
improve upon previously reported \ac{SoTA} results by a factor of 2-3.

\textbf{Chapter 6} consists of some concluding remarks. We highlight that the
inability of \ac{DL}-based systems to leverage established domain-specific 
principles is inexpedient. An approach to learning that acts within the
framework of laid-out rules and learns from them is a key objective of \ac{DL}
research. The findings in the dissertation are a step towards a general
framework for domain-obedient learning systems, paving the way for encouraging
future research opportunities. In particular, agentic systems reasoning toward
achieving goals could greatly benefit from domain-knowledge guidance, and
future work tackling this challenge would be compelling.

Domain knowledge-aware \ac{DL} also promises to be a major milestone towards
scientific applications of deep learning. In this vein, we undertook a
preliminary study attempting to address a problem arising from cosmology, where
the known laws of physics serve as constraining information. Results from this
study are presented in \textbf{Appendix A}.
