\chapter{Conclusion}
\label{chap:Conclusion}

There have been undeniably massive strides in the capabilities of modern deep
learning-based systems, as evidenced by their widespread adoption across several
fields. However, their inability to leverage established domain-specific 
principles is inexpedient. Furthermore, their tendency to violate these 
principles is a major hindrance to their adoption in critical fields like 
medicine, law, industrial automation, robotics, avionics, etc., where 
data-driven decision models must operate within domain-specific constraints. An 
approach to learning that acts within the framework of laid-out rules and 
leverages these rules alongside data to improve predictive performance is a key
objective of deep learning research.

This dissertation first analyzes deep learning-based systems with regard to
their ability to adhere to domain constraints when learning from data alone. A
persistent sentiment echoed by many researchers probing this area is the dearth
of large-scale rule-supported datasets; to address this issue, we put forth a
large-scale logical understanding evaluation dataset with rule annotations. This
dataset helps us demonstrate that large-scale training with an abundance of data
does not guarantee domain obedience. Similarly, intervention-based approaches
allow us to show that language models are not semantically faithful, i.e., they
do not behave in accordance with domain expectations. Following this, a new
technique to improve constraint adherence is presented, and issues with the
current paradigm of domain-blind evaluation of models are highlighted. Finally,
we explore domain coherence in the frontier of deep learning-based natural
language research--large language models--and propose techniques to improve
their domain obedience.

Domain knowledge-aware deep learning promises to be a major milestone towards
scientific applications of deep learning. In this vein, we undertook a
preliminary study attempting to address a problem arising from cosmology, where
the known laws of physics serve as constraining information. Results from this
study are presented in Appendix \ref{chap:LADDERastro}.

The findings in this dissertation are a step towards a general framework for
data-driven learning systems that operate within and learn from domain-specific
requirements, and they pave the way for encouraging future research
opportunities. In particular, agentic systems that reason toward achieving goals
could greatly benefit from domain-knowledge guidance, and further work addressing
this challenge would be compelling. Additionally, a natural extension of this
work is to couple constraint-aware models with contemporary explainability
techniques. Such integration would allow us to assess whether the model-generated
rationales align with the underlying domain rules and could facilitate the
creation of a transparent and consistent audit trail.

While the focus of this dissertation has been on advancing the applicability of 
deep learning-based systems in domain-specific contexts, it is also vital to 
consider the broader societal implications of these technological advancements. 
Although widespread use of automation technologies holds significant promise, it 
also threatens the economic futures of large swathes of the population in the 
current societal landscape. Policy shifts prioritizing strong social safety 
nets are crucial for mitigating these effects.

\insertdecoration
