\footnote{This appendix is largely based on our paper titled \emph{``LADDER:
Revisiting the Cosmic Distance Ladder with Deep Learning Approaches and
Exploring Its Applications''} \cite{myLadder}.}Constraint-aware \ac{DL}
techniques enable the application of \ac{DL} towards solving scientific
problems. In this Appendix, we explore such a problem arising from the field of
cosmology. In addition to constraints imposed by established laws of physics,
this problem presents extra hurdles, such as the samples not being independently
and identically distributed (i.i.d.) and the datasets being extremely limited in
size. We discuss the process involved in developing a first-of-its-kind
technique to tackle these challenges from the ground up and present empirical
results to validate our technique.

One of the core challenges in cosmology has been measuring distances. Several 
methods to indirectly estimate distances to far-flung objects have been devised,
each having its own region of operation. In addition, the aptly named 
\emph{cosmic distance ladder} principle allows us to calibrate the various
measurements using the potential regions of overlap between these disparate
measurement techniques.

In the conventional distance ladder approach \cite{Riess:2023egm}, we start with 
geometric measurements of nearby objects as the first ``rung'' of the ladder.
These measurements are then used to calibrate Cepheid variables
\cite{Freedman:2023zdo} or Tip of the Red Giant Branch (TRGB) stars
\cite{Freedman:2020dne}. Finally, using these measurements, we estimate
distances to the Type Ia supernovae (SNIa). On the other hand, the ``inverse''
distance ladder begins with the Cosmic Microwave Background (CMB), and through
various intermediate steps, ultimately extends to SNIa at lower redshifts 
\cite{Cuesta:2014asa,Camarena:2019rmj}. SNIa are favored as the final step for 
both ladders due to their reliability across a wide range of redshifts.\footnote{
\emph{Redshift} here refers to the frequency of the observed light being reduced 
owing to the object moving away from the observer as a consequence of the Doppler 
effect. Edwin Hubble first noted that there is a correlation between the distance 
to a stellar object and the speed at which it is moving away from the Earth. 
Given this well-established relationship, distances are often talked about in 
terms of the amount of redshifts.}

The current standard model of understanding of the evolution of the universe is 
$\Lambda$ Cold Dark Matter ($\Lambda CDM$), which has six free parameters that 
must be computed from observational data. However, employing this model 
alongside observational data has led to some inconsistencies--the most notable 
of which is the Hubble tension \cite{novosyadlyj, hazra, deg2}. This, alongside 
a growing list of other issues has prompted the cosmological community to device 
more complicated models or look for ``model-free'' approaches to the cosmic 
distance ladder paradigm. 

Alongside cosmography \cite{Visser:2004bf}, which utilizes a Taylor series,
several other methods like Gaussian processes, genetic algorithms, etc., have 
been studied in this context 
\cite{Keeley:2020aym, Mukherjee:2021kcu, Arjona:2020axn, Li:2023gpp}. However, 
ambiguity over polynomial degree, choice of kernels, convergence issues, 
overfitting concerns, and overwhelming errors in data-scarce regions, have 
hindered widespread adoption of these techniques 
\cite{OColgain:2021pyh, Hwang:2022hla}. Owing to these shortcoming, 
\ac{DL}-based approaches have gained prominence in this domain \citep{Wang:2019vxv, 
EscamillaRivera:2021vyw, Olvera:2021jlq, Gomez-Vargas:2022bsm, 
Gomez-Vargas:2021zyl, Giambagli:2023ngt, Dialektopoulos:2023dhb, 
Dialektopoulos:2023jam, Mukherjee:2024akt, Zhang:2023xgr, Zhang:2023ucf, 
Xie:2023ydk, Tang:2020nmw, Wang:2020hmn, Mehrabi:2023tld, Liu:2023rrr}. Despite 
several advances, \ac{DL}-based methods have not been able to extrapolate 
observational data to data-scarce regions in a consistent manner while taking 
into account all the available aleatoric noise information. To this end, we 
device a novel algorithm from first-principles called \textbf{LADDER} 
(\textbf{L}earning \textbf{A}lgorithm for \textbf{D}eep \textbf{D}istance 
\textbf{E}stimation and \textbf{R}econstruction) which is the first constraint 
aware \ac{DL}-based solution to this problem.

In the following sections we first formally state the problem and the associated 
constraints, followed by a presentation of our solution and elaborate on the design 
decisions. Finally, we present experimental results validating our approach.
